---
title: "Snakes & Ladders Analysis"
author: "Neto & Silva"
date: "March 2017"
output: 
  html_document:
    toc: true
    toc_depth: 2
---

```{r, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

Let's read the data with all the statistics:

```{r, message=FALSE}
read_csv2("../stats/game.stats.csv") %>% 
  mutate(int.rule = factor(int.rule)) %>%   # convert rule descriptions to factors
  mutate(end.rule = factor(end.rule)) -> d
```

```{r, echo=FALSE, eval=FALSE}
head(d)
```

## WIN criterion

The WIN criterion deals with how balanced the variants are, concerning winner percentages for each player.

```{r, fig.height=10}
d %>% 
  rowwise() %>% 
  mutate(win.sd = sd(c(winner1.perc, winner2.perc, winner3.perc,
                       winner4.perc, winner5.perc, winner6.perc), na.rm=TRUE)) %>% 
  select(game, n.players, int.rule, end.rule, win.sd) %>% 
  ggplot(aes(x=game, y=win.sd, shape=int.rule, colour=end.rule)) +
  geom_point(size=2.5) +
  scale_y_log10() +
  labs(x="game",y="standard deviation of winners", color="ending", shape="interaction",
       title="WIN criterion per number of players") +
  facet_wrap(~n.players) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

ggsave('WIN.pdf')
```

## WIN percentage


```{r, fig.height=8}
ideal.percentages <- data.frame(n.players=3:6, winner1.perc=1/3:6, int.rule="none")

d %>% 
  select(game, n.players, int.rule, end.rule, matches("winner")) %>% 
  ggplot(aes(x=n.players, y=winner1.perc, shape=int.rule, colour=end.rule)) +
  geom_line() +
  geom_point(size=2) +
  geom_line(data=ideal.percentages, color="black", linetype="dashed") +
  scale_y_log10() +
  labs(x="number of players",y="winning percentage for 1st player", 
       color="ending", shape="interaction") +
  facet_wrap(~game, scales = "free_y") +
  theme_bw()
```

```{r, fig.height=8}
ideal.percentages <- data.frame(n.players=3:6, winner3.perc=1/3:6, int.rule="none")

d %>% 
  select(game, n.players, int.rule, end.rule, matches("winner")) %>% 
  ggplot(aes(x=n.players, y=winner3.perc, shape=int.rule, colour=end.rule)) +
  geom_line() +
  geom_point(size=2) +
  geom_line(data=ideal.percentages, color="black", linetype="dashed") +
  scale_y_log10() +
  labs(x="number of players",y="winning percentage for 3rd player", 
       color="ending", shape="interaction") +
  facet_wrap(~game, scales = "free_y") +
  theme_bw()
```

The 6th player can only be visualized for 6 player games, and so we lose one dimension:

```{r}
d %>% 
  dplyr::filter(!is.na(winner6.perc)) %>% 
  select(game, n.players, int.rule, end.rule, winner6.perc) %>% 
  ggplot(aes(x=game, y=winner6.perc, shape=int.rule, colour=end.rule)) +
  geom_point(size=2) +
  geom_hline(yintercept = 1/6, linetype="dashed") +
  labs(x="game",y="winning percentage for 6th player", 
       color="ending", shape="interaction") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```


## Testing the Uniformity Hypothesis

```{r}
library(knitr)
library(xtable)
```

```{r}
d %>% 
  select(game, n.players, int.rule, end.rule, not.random) %>% 
  dplyr::filter(not.random <= 0.95) %>% 
  arrange(not.random) %>% 
  kable()
```

```{r, results="asis"}
d %>% 
  select(game, n.players, int.rule, end.rule, not.random) %>% 
  dplyr::filter(not.random <= 0.95) %>% 
  arrange(not.random) %>% 
  xtable() %>% 
  print(comment = FALSE)
```


## LEAD criterion

LEAD values state that a game is more dramatic if the player that wins
did achieve the lead only a few turns before the end of the game.

We are just plotting the swap 6 player variants for presentation purposes (the different interaction rules and less number of players give similar results):

```{r}
d %>% 
  select(game, n.players, int.rule, end.rule, matches("rounds")) %>% 
  select(game, n.players, int.rule, end.rule, matches("persistence")) %>% 
  dplyr::filter(n.players==6) %>% 
  tidyr::gather("rounds", "n.games", 5:24) %>% 
  dplyr::mutate(rounds = as.numeric(replace(rounds, 1:1260, rep(1:20,each=63)))) %>% 
  dplyr::filter(int.rule=="swap") %>%  
  ggplot(aes(x=rounds, y=n.games, shape=int.rule, colour=end.rule)) +
  geom_line() +
  geom_point(size=1.5) +
  labs(x="turns ahead before winning",y="number of games", 
       color="ending", shape="interaction") +
  facet_wrap(~game) +
  theme_bw()
```

## ANP criterion

The ANP criterion focus on the number of different players that were ahead
during a game. The argument is that a game is more dramatic the more
players are able to take the lead.

For visualization purposes we just show for the goose ending rule:

```{r}
d %>% 
  dplyr::filter(n.players==6) %>% 
  select(game, int.rule, end.rule, matches("players.ahead")) %>% 
  tidyr::gather("ahead", "total", 4:9) %>% 
  dplyr::mutate(ahead = as.numeric(replace(ahead, 1:378, rep(1:6,each=63)))) %>% 
  dplyr::filter(end.rule=="goose") %>% 
  ggplot(aes(x=ahead, y=total, shape=int.rule, colour=end.rule)) +
  geom_line() +
  geom_point(size=1.5) +
  labs(x="number of players ahead",y="percentage of games", 
       color="ending", shape="interaction") +
  facet_wrap(~game) +
  theme_bw()
```

The Dayking variant shows the best results in these criterion.




The next diagram shows the number of rounds each player was ahead in the entire simulation (should be as balanced as possible):

```{r}
d %>% 
  dplyr::filter(n.players==6) %>% 
  select(game, int.rule, end.rule, matches("ahead")) %>%
  select(game, int.rule, end.rule, matches("mean"), matches("sd")) -> d2

d2 %>% 
  select(-matches("sd")) %>% 
  tidyr::gather("ahead", "mean", 4:9) %>% 
  dplyr::mutate(ahead = as.numeric(replace(ahead, 1:378, rep(1:6,each=63)))) -> d2a

d2 %>% 
  select(-matches("mean")) %>% 
  tidyr::gather("ahead", "sd", 4:9) %>% 
  dplyr::mutate(ahead = as.numeric(replace(ahead, 1:378, rep(1:6,each=63))))-> d2b

d2a %>% 
  dplyr::inner_join(d2b) %>% 
  dplyr::filter(end.rule == "stop") -> d3


```

We present just for one type of ending rule. This rule does not change the interaction during the game, and eases the visualization:

```{r}
ggplot(d3, aes(x=ahead, y=mean, shape=int.rule, colour=end.rule)) +
  geom_line() +
  geom_point(size=1.5) +
  # geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.1) +  
  labs(x="i-th player ahead",y="percentage of total rounds", 
       color="ending", shape="interaction") +
  facet_wrap(~game) +
  theme_bw()
```

Notice however that the standard deviation is high, usually with the similar value as the mean itself. 

## Game Length

```{r, fig.height=6}
d %>% 
  dplyr::filter(n.players==6) %>% 
  select(game, n.players, int.rule, end.rule, matches("moves")) %>%
  ggplot(aes(x=int.rule, y=moves.mean, fill=end.rule)) +
  geom_bar(stat = "identity") +
  labs(x="interaction rule",y="mean game size") +
  facet_wrap(~game) +
  theme_bw()
```


## Log-Normal fit


```{r, message=FALSE, warning=FALSE}
best.fit <- d[which.max(d$lognorm.AIC),]
  
filename <- paste0(best.fit$game,"_",best.fit$n.players,"p_",
                   best.fit$int.rule,"_",best.fit$end.rule,"_50000.csv")
df <- read_csv2(paste0("../raw_data/",filename))[-1]

print(paste(best.fit$game,":",best.fit$n.players, "players with",
            best.fit$int.rule,"&", best.fit$end.rule,"rules"))
```

```{r}
res <- df$n.rounds
hist(res, breaks=50, ylab=NA, xlab="number of turns", prob=T, yaxt='n',xlim=c(0,200),
     main="", sub="")
xs  <- seq(min(res), max(res), len=100)
ys  <- dlnorm(xs, best.fit$lognorm.mean, best.fit$lognorm.sd)
points(xs,ys,type="l", col="red", lwd=2)
```

```{r, message=FALSE, warning=FALSE}
worst.fit <- d[which.min(d$lognorm.AIC),]
  
filename <- paste0(worst.fit$game,"_",worst.fit$n.players,"p_",
                   worst.fit$int.rule,"_",worst.fit$end.rule,"_50000.csv")
df <- read_csv2(paste0("../raw_data/",filename))[-1]

print(paste(worst.fit$game,":",worst.fit$n.players, "players with",
            worst.fit$int.rule,"&", worst.fit$end.rule,"rules"))
```

```{r}
res <- df$n.rounds
hist(res, breaks=12, ylab=NA, xlab="number of turns", prob=T, yaxt='n',
     main="", sub="")
xs  <- seq(min(res), max(res), len=100)
ys  <- dlnorm(xs, worst.fit$lognorm.mean, worst.fit$lognorm.sd)
points(xs,ys,type="l", col="red", lwd=2)
```

```{r}
d %>% 
  select(game, n.players, int.rule, end.rule, matches("log")) %>% 
  dplyr::filter(int.rule=="swap") %>% 
  ggplot(aes(x=n.players, y=lognorm.AIC, shape=int.rule, colour=end.rule)) +
  geom_line() +
  geom_point(size=2) +
  labs(x="number of players",y="AIC", 
       color="ending", shape="interaction") +
  scale_y_log10() +
  facet_wrap(~game) +
  theme_bw()
```

The values decrease with the number of players. This means player interaction together with the special cells, makes the game less predictable wrt to the lognormal model fit.



```{r}
d %>% 
  select(game, n.players, int.rule, end.rule, matches("log")) %>% 
  # dplyr::filter(int.rule=="swap") %>% 
  ggplot(aes(x=n.players, y=lognorm.mean, shape=int.rule, colour=end.rule)) +
  geom_line() +
  geom_point(size=2) +
  labs(x="number of players",y="Mean of lognormal fit", 
       color="ending", shape="interaction") +
  scale_y_log10() +
  facet_wrap(~game) +
  theme_bw()
```

This well-behaved lines are evidence that we can estimate the average mean and variance of these variants with more players.

